% Template for progress reports
% For examples on how to use latex see, for example,
% http://www.mat.bham.ac.uk/R.W.Kaye/latex/examples.htm
%
% Andy Sayler
% Constantin Berzan

\documentclass[12pt]{article}

\usepackage{times}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{times}
\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{url}

\setlength {\topmargin} {0 mm}
\setlength {\headsep} {0 mm}
\setlength {\headheight} {0 in}
\setlength {\voffset} {0 mm}
\setlength {\oddsidemargin} {0 mm}
\setlength {\evensidemargin} {0 mm}
\setlength {\hoffset} {0 mm}
\setlength {\textwidth} {6.5 in}
\setlength {\textheight} {9 in}

\hypersetup{
    colorlinks=true
}

\begin{document}

\title{BBR Progress Report 006:\\ Autonomous 2-D Mapping of a Building Floor}
\author{Andy Sayler \& Constantin Berzan}
\maketitle

\begin{abstract}
In the past two weeks, we implemented and tested landmark extraction based on
the RANSAC algorithm. We implemented the prediction phase of the EKF, and part
of the correction phase. We have found inconsistencies in the tutorial we are
following, which we are now trying to reconcile. The next step is to finish
implementing the EKF, either by computing the equations ourselves, or by finding
another source.
\end{abstract}

\emph{The code for this project is located at
\url{http://github.com/cberzan/bbr-mapper}}

\section{Introduction}
SLAM is the problem of Simultaneous Localization and Mapping using a mobile
robot.  We aim to develop a proof-of-concept SLAM system adhering to the
behavior-based philosophy.  The robot will use odometry and laser range data to
navigate the first floor of Halligan, and produce an image representation of
the floor map.

\section{Project Summary}

We will be developing our robot using a schema (or possibly hybrid)
architecture. The robot will utilize schema based functions to manifest the
following behaviors:

\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Avoid obstacles
    \item Avoid local minima
    \item Seek new areas
    \item Utilize SLAM (laser + odometry) to deduce current location
    \item Utilize SLAM (laser + odometry) to generate persistent environmental
          map
\end{itemize}

We will utilize the ADE robotics environment to complete our implementation.
The first objective will be to create code capable of navigating our robot
through unfamiliar environments and exploring these environments to their full
potential without getting stuck or colliding with obstacles. 

The next goal will be to build a map of the environment and provide
localization abilities. This will be done by implementing a basic SLAM
system using laser and odomtery data.  We aim to develop a 2D floor plan map
using data from our SLAM system.\\

Should we complete the initial scope of this project ahead of schedule, we may
opt to pursue one or more of the following extensions:

\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Utilize vision data in SLAM system
    \item Utilize vision data in 2D map generation
    \item Compare performance of our SLAM system to Carmen
    \item Augment SLAM system with additional sensor packages (radio ranging,
          etc)
\end{itemize}


\section{Problems tackled}

Literature review:
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item SLAM for Dummies tutorial -- both
    \item Carmen -- Andy
    \item Vision-based SLAM -- Constantin
\end{itemize}

\noindent{Discussed and refined project scope:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Opted to implement SLAM ourselves
    \item Decided to use SLAM data as primary source for persistent 2D map data
\end{itemize}

\noindent{Logistical issues:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Decided on workflow and code layout for interfacing our code with ADE
    \item Figured out bootstrapping of ADE registry and simulator
    \item Wrote SSH scripts for easy remote operation of the robot
\end{itemize}

\noindent{Discussed Odometry Testing Setup:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Linear Distance Testing (i.e. Drive 10m, then measure actual
      distance and angle)
    \item Start/Stop Testing (i.e. drive 1m in 10cm steps issuing
      start and stop commands between each step to measure start/stop drift)
    \item Rotational Distance Testing (i.e. Turn 180 degrees, then measure
      actual angle.)
    \item Acceleration Test (i.e. drive 10m and use internal timer to log
      distance at each tenth second interval to compute robot dynamics)
\end{itemize}

\noindent{Created Map of Halligan for the Simulator:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item The map was auto-generated by extracting all rectangles from a SVG
          image, which was drawn by hand
    \item The svg-to-xml conversion script is reusable, and allows the map to be
          modified easily, using a graphical editor such as Inkscape
    \item The goal is to capture the complexity of the environment, while not
          necessarily drawing it to scale
\end{itemize}

\noindent{Odometry Testing:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Performed drive-forward test:
        \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Varied distance: 1 m, 2 m
            \item Varied velocity: 0.1 m/s, 0.25 m/s, 0.5 m/s
            \item Surface: lenolium in Halligan
        \end{itemize}
    \item Performed start-stop test:
        \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Move interval: 2 sec; Stop interval: 1 sec
            \item Distance: 2 m
            \item Velocity: 0.1 m/s
            \item Surface: lenolium in Halligan
        \end{itemize}
    \item Studied odometry data to understand the coordinate system of the
          robot
    \item Characterized the error in odometry data versus actual motion:
        \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Surprisingly, the errors were greater at smaller speeds
            \item At small speeds, there was a consistent drift to the left
                  (this effect disappeared at higher speeds)
            \item The stop-start test did not increase the error significantly
            \item Mean error at 0.1 m/s was about 10\%
            \item Mean error at 0.25 m/s was about 8\%
            \item Mean error at 0.5 m/s was about 1\%
        \end{itemize}
\end{itemize}

\noindent{Implementation Plan:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item The initial implementation is going to work in the ADE simulator,
          using beacons as landmarks. This will allow us to implement and test
          the EKF separately from the landmark detection algorithm.
    \item The next step will be extracting landmarks from laser data. This can
          also be tested in simulation.
    \item Output useful mapping data. Test it in the simulator.
    \item Test everything in the real world. Tweak Kalman filter.
\end{itemize}

\noindent{Odometry data in the simulator:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item We managed to obtain odometry data from ADESim
    \item We modified ADESim to add Gaussian noise to the odometry data on each
          update. We may decide to change the noise model later
\end{itemize}

\noindent{High-level Design:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
  \item Coded high level SLAM interfaces
      \begin{itemize}
          \item EKFServer - Provides best estimate of current location by
            performing EKF.
          \item LandmarkServer - Provides a list of currently seen landmarks
            and their location relative to the robot.
          \item MappingServer - Maintains persistent representation of map and
            outputs a map representation to the user.
          \item Arch - Implements exploratory schema based behavioral code.
      \end{itemize}
    \item Wrote config file to start all necessary servers.
    \item Learned to pass custom objects via java RMI.
\end{itemize}

\noindent{SLAM implementation:}
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Landmark extraction
    \begin{itemize}
        \setlength{\itemsep}{0pt}
        \setlength{\parskip}{0pt}
        \setlength{\parsep}{0pt}
        \item Fit lines to laser readings using RANSAC.
        \item Use estimated robot position from the EKF to find origin of
              global coordinate system.
        \item Project origin of coordinate system onto each fitted line, and
              take the projected point as a global point landmark.
        \item Report landmarks to the EKF only if they have been reobserved a
              sufficient number of times.
        \item ``Forget'' landmarks that have not been reobserved in a long time.
        \item Wrote a simple architecture to help find the distance between the
              robot's center of rotation and the position of the LRF.
        \item Added correction for the distance between the robot's center of
              rotation and the position of the LRF \emph{in simulation}. The
              correction constant will be different on the real robot.
        \item Testing: offline visualization of RANSAC results, stepping through
              the iterations of the algorithm.
        \item Testing: online visualization of RANSAC results, written on top of
              LRFServerVis.
        \item Testing: in simulation, the robot consistently re-observes
              landmarks when driving and turning in a random manner.
    \end{itemize}
    
    \item EKF
    \begin{itemize}
        \setlength{\itemsep}{0pt}
        \setlength{\parskip}{0pt}
        \setlength{\parsep}{0pt}
        \item Prediction
          \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Update location from delta odometry. Use
              delta odometry to minimize odometry error
              (odometry most trusted for small deltas) - Done
            \item Update covariance with new robot prediction. Update
              robot location covariance and robot to feature covariance. - Done
          \end{itemize}
        \item Correction
          \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Cycle through all old landmarks. Calculate Kalman Gain,
              predicted location, and delta location for each. Use each
              to update robot and landmark position.
            \item Update covariance with landmark position deltas and Kalman
              gains.
            \item Ignore old landmarks that have not been re-observed.
            \item Add new landmarks to state and covariance matrix.
          \end{itemize}
    \end{itemize}
\end{itemize}

\section{Next Steps}

\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parskip}{0pt}
    \setlength{\parsep}{0pt}
    \item Finish EKF Implementation -- Andy
      \begin{itemize}
            \setlength{\itemsep}{0pt}
            \setlength{\parskip}{0pt}
            \setlength{\parsep}{0pt}
            \item Resolve inconsistencies in ``Slam for Dummies'' tutorial.
            \item Add covariance update to EKF correction step
      \end{itemize}
    \item Test and tune EKF Implementation -- Andy
    \item Create contingency plan if EKF fails to operate -- Both
    \item Figure out persistent mapping -- Constantin
    \item Create visualization of persistent mapping for the demo -- Constantin
\end{itemize}

\end{document}
